### 2016-07-04 doc/publications/plos16/plos16.tex
changes to plos.tex

While unsupervised clustering is useful for selected clinical applications~\cite{Li2015}, a more common analysis task is
supervised learning, i.e. given a phenotype of clinical label determine the most associated features.  One of the
simplest supervised machine learning methods is logistic regression (LR), which creates a decision boundary by fitting a
linear combination of the feature variables.  Processing the `wide' genomic data poses difficulties for machine learning
applications as having more features $p$ than samples $n$ causes the model to overfit the data and generalize poorly.
This is known as the `curse of dimensionality'~\cite{Bauer2014}.  Random Forests~\cite{Breiman.2001} (RF) 
apply the technique of bootstrap aggregation or bagging to decision trees. The training set is randomly sampled $B$
times with replacement. A tree is grown on each of these $B$ samples and the results are combined
\begin{equation*}
{\bar {h}}={\frac {1}{B}}\sum _{b=1}^{B} h(x;\theta_b)
\end{equation*}

>>>>

While unsupervised clustering is useful for selected clinical applications~\cite{Li2015}, a more common analysis task is
supervised learning, i.e. given a phenotype of clinical label determine the most associated features.  One of the
simplest supervised machine learning methods is logistic regression (LR), which creates a decision boundary by fitting a
linear combination of the feature variables.  Processing the `wide' genomic data poses difficulties for machine learning
applications as having more features $p$ than samples $n$ causes the model to overfit the data and generalize poorly.
This is known as the `curse of dimensionality'~\cite{Bauer2014}.

A random forest is an ensemble  of decision $h(x; \theta_b) b=1, \ldots, B$ where
$x$ is the covariate vector of length $p$ with associated
random vector $X$ and the $\theta_b$ are i.i.d. random vectors. 
The training  data is independently drawn from the joint distribution of $(X,Y)
and comprises $n$  $(p+1)$-tuples (x_1,y_1),\ldots, (x_n,y_n)$.
For regression, the random forest prediction is the unweighted average over
the collection: $\bar{h}(x)=\dfrac{1}{K}\Sigma_{k=1}^K h(x; \theta_k)$.

The average prediction error for an individual tree $h(X; \theta)$ as
\begin{equation}
PE_t E_\theta E_{X,Y} (Y-h(X; \theta))^2
\end{equation}
Assume that for all  $\theta$ the tree is unbiased, i.e., $EY=E_Xh(X; \theta)$. Then
\begin{equation}
PE_f = \rho PE_t
\end{equation}
where $\rho$ is the weighted correlation between residuals
$(Y-h(X;\theta)$ and  $(Y-h(X;\theta))$
for independent $\theta, theta^\prime$.
The inequality (3) pinpoints what is required for accurate random forest regression:
(i) low correlation between residuals of differing tree members of
the forest, and (ii) low prediction error for the individual trees. Further, the
random forest will, in expectation, decrease the individual tree error, $PE_t$ , by
the factor $\rho$. Accordingly, the randomization injected strives for low correlation.
The strategy employed to achieve these ends is as follows:
1. To keep individual error low, grow trees to maximum depth.
bench.rf.

Random Forests~\cite{Breiman.2001} (RF) 
apply the technique of bootstrap aggregation or bagging to decision trees. The training set is randomly sampled $B$
times with replacement. A tree is grown on each of these $B$ samples and the results are combined
\begin{equation*}
{\bar {h}}={\frac {1}{B}}\sum _{b=1}^{B} h(x;\theta_b)
\end{equation*}
